### 2019-02-22 frontmatter.tex
Update front matter and title etc

%-------------------------------------------------------------------------------
% Cover page
%-------------------------------------------------------------------------------
%
\makecoverpage

%-------------------------------------------------------------------------------
% Dedication (comment if none)
%-------------------------------------------------------------------------------
%
\dedication{
	{\it ``You must unlearn what you have learnt.''}\\[20pt]
	Master Yoda
}

%-------------------------------------------------------------------------------
% Abstract (English)
%-------------------------------------------------------------------------------
%
% COMMENT: the abstract should be limited to 2000 characters and the keywords
% to 250 characters because otherwise they will not fit in the mailing sheet.
%
% TODO: abstract in english and swedish
\begin{abstract}
        To do
	% Access points  must work. In fact, few cryptographers would disagree with
	% the emulation of robots, which embodies the technical principles of
	% complexity theory. Hyp, our new application for lossless information, is the
	% solution to all of these challenges.
	%         %
	%         \keywords{super-lasers, space-stations, The Force, cookies.}
	%         %
\end{abstract}

%-------------------------------------------------------------------------------
% Abstrakt (Swedish) 
%-------------------------------------------------------------------------------
%
% COMMENT: the abstract should be limited to 2000 characters and the keywords
% to 250 characters because otherwise they will not fit in the mailing sheet.
%
\begin{abstrakt}
        To do
	% \r{A}tkomstpunkter m\r{a}ste arbeta. I sj\"{a}lva verket skulle f\r{a}
	% kryptografer h\r{a}ller inte med emulering av robotar, som
	% f\"{o}rkroppsligar de tekniska principer av komplexitetsteori. Hyp, v\r{a}r
	% nya applikation f\"{o}r f\"{o}rlustfri information är l\"{o}sningen på alla
	% dessa utmaningar.
	%         %
	%         \nyckelord{super-lasrar, rymdstationer, Kraften, kakor.}
	%         %
\end{abstrakt}

%-------------------------------------------------------------------------------
% Preface page
%-------------------------------------------------------------------------------
%
\begin{preface}
        % TODO: preface
        To do.
	% This thesis deals with high energy super-lasers. A brief introduction on the
	% basic concepts and methods is presented in the first part. The second part
	% contains two articles.
	% The papers are adjusted to comply with the present thesis format for
	% consistency, but their contents have not been altered as compared with their
	% original counterparts.
\end{preface}

%-------------------------------------------------------------------------------
% Division of work between authors
%-------------------------------------------------------------------------------
%
%TODO: division of work
\begin{divisionofwork}
	The main advisor for the project is Erik Lindborg (EL).
	Pierre Augier (PA) acts as co-advisor.

        To do: paper wise list
	% \paperitem
	%         The code has been developed by Anakin Skywalker (AS).
	%         The paper has been written by AS and Darth Vader (DV).
        %
	% \paperitem
	%         The experimental set-up has been designed by DS.
	%         The simulations have been performed by AS using the control-code
	%         developed by DS.
	%         The paper has been written by AS and DV with feedback from OK and DS.

\end{divisionofwork}

%TODO: fill in additional publications and conferences
%-------------------------------------------------------------------------------
% Additional publications (comment if none)
%-------------------------------------------------------------------------------
%
% \begin{otherpublications}
%         The following papers, although related, are not included in this thesis.
%
%   \paperitem%
%     {Anakin Skywalker \& Master Obi-Wan Kenobi}% Authors
%     {3639}% Year
%     {The light sabre: an elegant weapon for a more civilized age}% Title
%     {Jedi Journal of Weapons}% Journal
%     {33}% Volume
%     {2}% Number
%     {pp. 55--60}% Pages
%
% \end{otherpublications}

%-------------------------------------------------------------------------------
% Conferences (comment if none)
%-------------------------------------------------------------------------------
%
% \begin{conferences}
%         Part of the work in this thesis has been presented at the following
%         international conferences. The presenting author is underlined.
%
%   \conferenceitem%
%     {Anakin Skywalker, \underline{Darth Vader} \& Admiral Motti}% Authors
%     {3639}% Year
%     {Effects on breathing of underestimation of the power of The Force}% Title
%     {Intergalactic Conference on Weapons and Peace}% Conference
%     {Kamino}% Location
%
% \end{conferences}

%-------------------------------------------------------------------------------
% Table of contents
%-------------------------------------------------------------------------------
%
\tableofcontents