### 2019-05-31 paper_05_milestone_issf/table.tex

\begin{table}
\caption{Parameters explored in our experiments without system rotation.  The
line written in bold font indicates the experiment described in the last
section of the present paper. $F_{hc}$ and $\R_c$ are based on the carriage
velocity $U_c$ and the mesh $M$ as defined in equations (\ref{eq:c}).  $F_{he}$
and $\R_e$ are estimations of the turbulent Froude and buoyancy Reynolds
numbers as defined in equations (\ref{eq:FhR}) and (\ref{eq:e}), i.e.\ with the
estimations $E_K \simeq 0.08 U_c^2$ and $\epsK \simeq 0.02 {U_c}^3/M$.}
\centerline{\begin{tabular}{| c | c | r | r | r | r |}
\hline
{$N$~(rad/s)} & {$U_c$~(cm/s)} & {\bf $F_{hc}$} & {\bf $\R_c$} &
{\bf $F_{he}$} & {\bf $\R_e$} \\
\hline
0 & [2, 4, 6,  &  $\infty$ & $\infty$ &  $\infty$ & $\infty$  \\
&  8, 12, 16] & ~~~~~~~~~~ & ~~~~~~~~~~ & ~~~~~~~~~~ & ~~~~~~~~~~\\
\hline
    &    2 & 0.07 &    67 &  0.02 &   1.3 \\ \cline{2-6}
    &    4 & 0.13 &   533 &  0.03 &  10.7 \\ \cline{2-6}
0.4 &    6 & 0.20 &  1800 &  0.05 &  36.0 \\ \cline{2-6}
    &    8 & 0.27 &  4267 &  0.07 &  85.3 \\ \cline{2-6}
    &   12 & 0.40 & 14400 &  0.10 & 288.0 \\ \cline{2-6}
    &   16 & 0.53 & 34133 &  0.13 & 682.7 \\
\hline
    &    2 & 0.03 &    17 &  0.01 &   0.3 \\ \cline{2-6}
    &    4 & 0.07 &   133 &  0.02 &   2.7 \\ \cline{2-6}
0.8 & {\bf 6.00} & {\bf 0.10} & {\bf  450} & {\bf  0.02} & {\bf  9.0} \\ \cline{2-6}
    &    8 & 0.13 &  1067 &  0.03 &  21.3 \\ \cline{2-6}
    &   12 & 0.20 &  3600 &  0.05 &  72.0 \\ \cline{2-6}
    &   16 & 0.27 &  8533 &  0.07 & 170.7 \\
\hline
\end{tabular}}
\label{table}
\end{table}