### 2019-06-05 chapter_01_decomposition.tex
Section on helmholtz normal mode decomposition + glossary

\section{Helmholtz and normal mode decomposition}

\subsection{Governing Equations}

The governing equations for a shallow layer of fluid are:
\begin{align}
  \partial_t \mathbf u 
      = &- (\mathbf{u}.\nabla) \mathbf{u} - c^2 \nabla h - f\mathbf{e_z} \times 
\mathbf u \label{dtu0} \\
  \partial_t h 
      =& - \nabla. (h \mathbf u) \label{dth}
\end{align}
Equation \eqref{dtu0} may also be written in the \textit{rotational form} as:
\begin{equation}
 \label{dtu}
 \partial_t \mathbf u 
 = - \nabla |\mathbf u|^2/2 - c^2 \nabla h - \zeta \times \mathbf u
\end{equation}
where, \gls{zeta}
represents absolute vorticity, i.e. the sum of vorticity and system rotation.
Furthermore, equations \eqref{dtu0} and \eqref{dth} can be combined to form an 
equation for the \textit{total mass flux}, \gls{J} $= h\mathbf{u}$:
\begin{equation}
  \label{dtJ}
 \partial_t \mathbf J = -(\mathbf{u}.\nabla)\mathbf{J} - \nabla(c^2h^2)/2 -
\zeta \times \mathbf J - (\nabla. \mathbf J)\mathbf u
\end{equation}
Similarly, an equation for the \textit{displaced mass flux}, \gls{M} $ = 
\eta\mathbf{u}$ is:
\begin{equation}
  \label{dtM}
 \partial_t \mathbf M = -(\mathbf{u}.\nabla)\mathbf{M} - \nabla(c^2\eta^2)/2 - 
\zeta \times \mathbf M - (\nabla.\mathbf{u} + \nabla.\mathbf{M})\mathbf u
\end{equation}
For a divergence free flow, a Poisson equation for \emph{h} can be formulated.
Taking divergence of \eqref{dtu},
yields the Poisson equation:
\begin{equation}
  \label{poisson}
  \nabla^2 h = \frac{1}{c^2} \left[ \nabla.(\zeta \times \mathbf u )
	      - \nabla^2 \frac{|u|^2}{2} \right]
\end{equation}
Applying fourier transform, the spectral counterpart for \eqref{poisson} in 
tensor notation is:
\begin{equation}
\label{poisson_fft}
  -\kappa^2 \hat{h} = \frac{1}{c^2} \left[ ik_i (\widehat{\epsilon_{ijk} \zeta_j 
u_k})
	      + \kappa^2 \frac{\widehat{u_i u_i}}{2} \right]
\end{equation}

\subsection{Helmholtz Decomposition}
\subsubsection{For velocity field, u}
The Helmholtz decomposition theorem, or the fundamental theorem of vector 
calculus, 
states that any well-behaved vector field can be decomposed into
the sum of a longitudinal (diverging, non-curling, irrotational) vector field 
and 
a transverse (solenoidal, curling, rotational, non-diverging) vector field. 
This allows us to express the velocity field as:
\begin{align}
\label{helm_u}
\mathbf u & = -\nabla \times (e_z \Psi) + \nabla \Phi\\
	& =  -\nabla \times \Psi_z + \nabla \Phi
\end{align}

For the sake of clarity,
we shall denote the rotational (vortical) and divergent (wave) parts of the 
velocity with the suffix \emph{r} and \emph{d} respectively. Thus,
\begin{align}
\mathbf u_r & = -\nabla \times \Psi_z ; &\mathbf u_d =   \nabla \Phi
\end{align}
And therefore, $\mathbf u  = \mathbf u_r + \mathbf u_d$.

To find the projection operators for the divergent part, taking divergence of 
\eqref{helm_u} gives $\nabla .\mathbf{u} = \nabla^2 \Phi
$.
This transforms in the spectral plane as, $ik_j \hat{u}_j = -\kappa^2 
\hat{\Phi}$, implying:
\begin{align}
\mathbf{\hat{u}}_d =& ik_i \hat{\Phi} = \frac{k_i k_j}{\kappa^2} \hat{u}_j\\
\mathbf{\hat{u}}_r =& \mathbf{\hat{u}} - \mathbf{\hat{u}}_d = \left( \delta_{ij} 
- \frac{k_i k_j}{\kappa^2} \right) \hat{u}_j
\end{align}

Thus, for two-dimensions the decomposed velocity are represented as follows,

\begin{align}
\mathbf{\hat{u}}_d^x =& \frac{k_x (k_x \hat{u}_x + k_y \hat{u}_y)}{\kappa^2} \\
\mathbf{\hat{u}}_r =& \hat{u}_{x} - \frac{k_x  (k_x \hat{u}_x + k_y \hat{u}_y)}{\kappa^2} 
\end{align}

\subsubsection{For fluid depth, h}
To obtain a similar decomposition, $h = h_r + h_d$, one should use the Poisson 
equation \eqref{poisson_fft}.
Since Poisson equation requires a divergence free flow, the LHS of 
\eqref{poisson_fft} would correspond to the rotational part of the flow in the 
transformed plane, i.e. $\hat{h}_r$. Henceforth, the divergent part, $\hat{h}_d$ 
can be obtained by subtracting $\hat{h}_r$ from $\hat{h}$.

\subsection{Normal mode decomposition}

\subsubsection{Following Bartello 1995}
In order to isolate oscillating fast modes that display no nonlinearity, one 
can adopt a linearization approach followed by a normal mode decomposition. 
As a result of linearization, from \eqref{dtu} and \eqref{dth}:
  \begin{align}
    \partial_t \mathbf u =& - c^2 \nabla \eta - f\mathbf{e_z} \times \mathbf u 
  \label{dtu_l} \\
    \partial_t \eta =& - \nabla.  \mathbf u \label{dth_l}
  \end{align}
Taking curl and divergence of the linearized equation for momentum 
\eqref{dtu_l} gives the following evolution equations with change of variables:
  \begin{align}
    \partial_t \zeta =& - f \delta
  \label{dtcurl_l} \\ 
    \partial_t \delta =&  f \zeta - c^2 \nabla^2 \eta \label{dtdiv_l} \\
    \partial_t \eta =& - \delta \label{dteta_l}
  \end{align}
where $\zeta$ and $\delta$ are relative vorticity and divergence as functions 
of $\mathbf{r}$ and $t$ respectively. Representing the dependent flow 
quantities in terms of Fourier modes:
  \begin{align}
  \begin{pmatrix}
      u \\ v \\ \eta \\ \zeta \\ \delta
  \end{pmatrix} (\mathbf{r},t)
  = \int  
  \begin{pmatrix}
      \hat{u} \\ \hat{v} \\ \hat{\eta} \\ \hat{\zeta} \\ \hat{\delta}
  \end{pmatrix} e^{i(\mathbf k . \mathbf{r} - \omega t)} \mathbf{dk} d\omega
  \end{align}
allows to rewrite the system of equations \eqref{dtcurl_l} to \eqref{dteta_l} 
as an eigenvalue problem:
\begin{align}
 i\omega
   \begin{Bmatrix}
       \hat{\zeta} \\ \hat{\delta} \\c\kappa\hat{\eta}
  \end{Bmatrix}
  = i
  \begin{bmatrix}
     0 & if & 0 \\
   -if &  0 & -ic\kappa \\
     0 & ic\kappa & 0 
  \end{bmatrix}
  \begin{Bmatrix}
       \hat{\zeta} \\ \hat{\delta} \\c\kappa\hat{\eta}
  \end{Bmatrix}
\end{align}
where, $\kappa = |\mathbf{k}|$.

Let us define $A$ as the Hermitian matrix operating on the vector 
$\mathbf{W} = \{\hat{\zeta}, \hat{\delta} ,c\kappa \hat{\eta} \}^T $.which 
yields the 
familiar dispersion relation 
for the slow geostrophic mode 
and fast Poincar\'e wave modes:
\begin{equation}
 \omega^{(0)} = 0;\quad \omega^{(\pm)}=\pm \sigma
\end{equation}
where, $\sigma = \sqrt{f^2 + c^2\kappa^2 }$. These are the eigenvalues of 
the matrix operator \emph{A}. Since \emph{A} is Hermitian, the corresponding 
eigenvectors are orthogonal and these are normalized as follows
\begin{align}
  \mathbf X^{(0)}_n = 
    \frac{1}{\sigma}
    \begin{Bmatrix}
       -c\kappa \\ 0 \\ f
     \end{Bmatrix}; \quad
  \mathbf X^{(\pm)}_n = 
    \frac{1}{\sqrt{2} \sigma}
    \begin{Bmatrix}
       f \\ \mp i\sigma \\ c\kappa
     \end{Bmatrix}
\end{align}
Let $X_n$ be the eigenvector matrix, which follows the 
property $X_n \bar{X}_n^{T}=I$. It can, then, be applied to diagonalize the 
system of 
equations as follows:
\begin{align}
 \partial_t \mathbf{W} =& [A] \mathbf{W}\\
 \partial_t (\bar{X}_n^T \mathbf{W}) =& \bar{X}_n^T[A]X_n 
\bar{X}_n^T\mathbf{W}\\
			    =& [\Lambda] (\bar{X}_n^T\mathbf{W})
\end{align}
Thus, the alternate diagonalized system of equations for the normal modes are 
given by:
\begin{align}
 \partial_t
\mathbf{N}
  = &
  \begin{bmatrix}
     0 & 0 & 0 \\
     0 & -i\sigma & 0 \\
     0 & 0 & i\sigma 
  \end{bmatrix}
  \mathbf{N}\\
\text{where},
   \mathbf{N} = \bar{X}_n^T \mathbf{W}
   = &\frac{1}{\sqrt{2}\sigma}
   \begin{Bmatrix}
       -\sqrt{2}c\kappa(\hat{\zeta} -f \hat{\eta}) \\ 
        f\hat{\zeta} + c^2\kappa^2\hat{\eta} - i\sigma\hat{\delta} \\
       f\hat{\zeta} + c^2\kappa^2\hat{\eta} + i\sigma\hat{\delta}
   \end{Bmatrix}
\end{align}

The first mode represents \emph{linearised potential vorticity} which is 
conserved in this framework. The remaining two are the \emph{linearised 
ageostrophic or wave modes}. In the absence of system rotation, i.e. $f=0$, 
the normal modes are simply,
\begin{equation}
\label{nmode}
 \mathbf{N} = 
 \begin{Bmatrix}
  \hat \zeta \\
  \frac{1}{\sqrt{2}} (c\kappa \hat{\eta} - i\hat{\delta})\\
  \frac{1}{\sqrt{2}} (c\kappa \hat{\eta} + i\hat{\delta})
 \end{Bmatrix}
\end{equation}

%---------------------------------------------%
\subsubsection{Following Farge \& Sadourny 1989}
Instead of finding the normal modes for the vorticity, divergence and 
displacement field of the flow, we shall make use of the Helmholtz 
decomposition described in \eqref{helm_u}. The shallow water equations then 
transform to:
 \begin{align}
    \partial_t \psi =&  f \phi
  \label{dtpsi_l} \\ 
    \partial_t \phi =&  -f \psi - c^2 \eta \label{dtphi_l} \\
    \partial_t \eta =& - \nabla^2 \phi \label{dteta_l2}
  \end{align}
where $\psi$ and $\psi$ are stream function and velocity potential as functions 
of $\mathbf{r}$ and $t$ respectively. By substituting the dependent 
variables with the respective Fourier transform, this reduces to the eigenvalue 
problem:
\begin{align}
 i\omega
   \begin{Bmatrix}
       \hat{\psi} \\ \hat{\phi} \\ \hat{\eta}
  \end{Bmatrix}
  = i
  \begin{bmatrix}
     0 & if & 0 \\
   -if &  0 & ic^2 \\
     0 & -i\kappa^2 & 0 
  \end{bmatrix}
  \begin{Bmatrix}
       \hat{\psi} \\ \hat{\phi} \\ \hat{\eta}
  \end{Bmatrix}
\end{align}
the square matrix is not Hermitian and this would result in complex 
eigenvalues. By adopting the following change of variables:
\begin{equation}
 \hat{\psi} \to \kappa^2\hat \psi = \hat{\zeta} ; \quad
 \hat{\phi} \to -\kappa^2\hat \phi = \hat{\delta}; \quad
 \hat{\eta} \to c\kappa\hat \eta
\end{equation}
it falls back to the previous eigenvalue problem as demonstrated in the  
previous section. In other words, we can use the same eigenvector matrix, $X_n$ 
to find the normal modes of:

$$\mathbf{H} = \{\hat \psi,\; \hat \phi,\;  \eta  \}^T$$
which is closely related to:
$$\mathbf{W} 
= \{\kappa^2\hat \psi,\; -\kappa^2\hat 
\phi,\; c\kappa\hat \eta  \}^T 
= \{\hat \zeta;\; \hat \delta;\;  c\kappa \hat \eta \}^T $$

%---------------------------------------------%
\subsubsection{Normal mode inversion to primitive variables}
To begin with, we shall form a new vector $\mathbf{B} = \mathbf{N}/\kappa$ 
which will have the same dimensions as velocity.
It has been shown before that the normal modes can be represented by $\mathbf{N} 
= \bar{X}_n^T \mathbf{W}$. Now,  we shall form a new vector $\mathbf{B} = 
\mathbf{N}/\kappa$ which will have the same dimensions as velocity. Thus,
\begin{align}
   \mathbf{B}
   = &\frac{1}{\sqrt{2}\sigma}
    \begin{Bmatrix} \sqrt{2} c\left(- 
    \kappa^{2} \hat \psi +  f\hat\eta \right)\\
    \kappa \left(c^{2} \eta + f \psi + i \phi \sigma\right)\\
    \kappa \left(c^{2} \eta + f \psi - i \phi \sigma\right)
    \end{Bmatrix}
\end{align}

The $\mathbf{B}$ vector can also be related to the primitive variable vector, 
$\mathbf{U} = \{\hat{u},\hat{v},c\hat{\eta}\}^T$ using transformation matrices 
as follows:
\begin{align}
 \mathbf{B} =&\frac{1}{\kappa}[\bar{X}_n]^T W\\
	    =&\frac{1}{\kappa}[\bar{X}_n]^T [P] \mathbf{U} 
\end{align}
where,
\begin{equation}P=
\begin{bmatrix}- i k_{y} &  i k_{x} & 0\\ i k_{x} &  i 
k_{y} & 
0\\0 & 0 & \kappa\end{bmatrix}
\end{equation}
It is also straightforward to show that the magnitude of the normal modes 
equals the total energy as:
\begin{align}
 \mathbf{\bar B}^T\mathbf{B} 
 =&\frac{1}{\kappa^2} \mathbf{\bar U}^T 
     [\bar P]^T[{X}_n]  [\bar{X}_n]^T [P] \mathbf{U} \\
 =& \mathbf{\bar U}^T\mathbf{U}
\end{align}
implying, $$2(E_K+E_P)=\sum_i B^{(i)}\bar{B}^{(i)}=\kappa^{2} \psi 
\overline{\psi} +\kappa^{2} \phi \overline{\phi} + c^{2} 
\eta \overline{\eta}
= u\bar{u} + v\bar{v} + c^2\eta\bar{\eta}$$
This allows us to represent the primitive variables in terms of normal modes as 
$$\mathbf{U} = [Q]\mathbf{B}$$
where, the inversion matrix is
\begin{align}
 Q 
 = & \kappa [P]^{-1}[\bar{X}_n]^{T^{-1}}\\
 =& \frac{1}{\sqrt{2}\sigma\kappa}
 \begin{bmatrix}
      -i\sqrt{2}c\kappa  k_{y} & 
      \left(i f k_{y} + 
	  k_{x}   \sigma\right) &
      \left(i f k_{y} - k_{x} \sigma\right)\\
      i \sqrt{2}c\kappa k_{x} & 
      \left(-  i f k_{x} +  k_{y}   \sigma\right) &
      - \left(i f  k_{x} + k_{y} \sigma\right)\\
      \sqrt{2}\kappa f & 
       c\kappa^{2} &
      c\kappa^{2}
\end{bmatrix}
\end{align}
In tensor notation,
\begin{align}
 \hat{u}_l =& \epsilon_{lm3} ik_m\left[ -\frac{c}{\sigma} B^{(0)} + 
	      \frac{f}{\sqrt{2}\sigma\kappa} (B^{(+)} + B^{(-)})
	      \right]
	      + k_l \frac{1}{\sqrt{2}\kappa}  (B^{(+)} - B^{(-)})\\
c\hat{\eta} =& \frac{f}{\sigma} B^{(0)} + \frac{c\kappa}{\sqrt{2}\sigma} 
(B^{(+)} + B^{(-)})    
\end{align}

\subsection{Spectral energy budget}
\subsubsection{Kinetic energy}
%---------------------------------------------%
\paragraphbf{Expression for mean kinetic energy}
Kinetic energy being cubic for shallow water equations, it can be split into 
quadratic and non-quadratic parts as follows:
\begin{align*}
  E_K(\mathbf{r},t)
    =& h\mathbf{u}.\mathbf{u}\\
    =& (1+\eta)\mathbf{u}.\mathbf{u}\\
    =& \mathbf{u}.\mathbf{u} + \eta\mathbf{u}.\mathbf{u}\\
    =& E_K^Q + E_K^{NQ}
\end{align*}

%---------------------------------------------%
\paragraphbf{K.E. budget}
Let the rate of change of kinetic energy, without any approximations, be 
written as:
\begin{equation}\label{dtKE}
 \pder{t}E_K(\mathbf{k},t) = T_K + C_K
\end{equation}
where, $T_K$ and $C_K$ represents the transfer and conversion spectral 
functions respectively. Equation 
\eqref{dtKE} splits into:
\begin{align*}
 \pder{t}E_K^Q + \pder{t}E_K^{NQ}
 =&  (T_K^{Q} + C_K^{Q}) + (T_K^{NQ} + C_K^{NQ})
%  =& \frac{1}{2}\partial_t(\mathbf{\hat u}.\widehat{h\mathbf{u}}^*)\\
%  =& \frac{1}{2}\partial_t(\mathbf{\hat u}.\hat{\mathbf{J}}^*)
\end{align*}
%-------------------------------------------------%
\paragraph{Quadratic K.E. budget:}
Consider the rate of change of quadratic kinetic energy,
\begin{equation}\label{dtKE2}
 \pder{t}E_K^{Q}(\mathbf{k},t) = T_K^{Q} + C_K^{Q}
\end{equation}
Following \eqref{dtu0}, equation \eqref{dtKE2} expands into:
\begin{align*}
 \partial_t E_K^{Q}(\mathbf{k},t)
 =& \frac{1}{2}\partial_t(\mathbf{\hat u}.\mathbf{\hat u^*})\\
 =& \frac{1}{2}\left[ \mathbf{\hat u} .\pder[\mathbf{\hat u^*}]{t} 
		      + \mathbf{\hat u^*}. \pder[\mathbf{\hat u}]{t}\right]\\
 =& \frac{1}{2}\left[ -\hat{u}_i (\widehat{ u_j\partial_j u_i })^* 
    - c^2 \hat{u}_i (ik_i \hat{\eta})^* 
    - \hat{u}_i (\epsilon_{i3k} f \hat{u}_k)^*
    - ... \text{conjugate terms}
    \right]\\
 =& -\text{Re}\left[ \hat{u}_i (\widehat{ u_j\partial_j u_i })^* 
    + c^2 \hat{u}_i (ik_i \hat{\eta})^* 
    + \hat{u}_i (\epsilon_{i3k} f \hat{u}_k)^* \right]
\end{align*}
% since, $u_i (\epsilon_{ijk} \zeta_j u_k) = 0$
% \footnote{ \textbf{Note}
% $\epsilon_{ijk} \zeta_j u_k 
% = \epsilon_{jki}  (\epsilon_{jlm} \partial_l u_m +   f_j )u_k$}
% =& (\delta_{kl}\delta_{im} - \delta_{km} \delta_{il})\partial_l u_m +    
% \epsilon_{ijk} f_j u_k\\
% =& u_k\pder{x_k}u_i - \pder{x_i}u_ku_k + \epsilon_{ijk} f_j u_k
we have,
\begin{align}
 T_K^{Q}= &-\text{Re}\left[\hat{u}_i (\widehat{ u_j\partial_j u_i })^* 
	   \right]\\
 C_K^{Q}= &-\text{Re}\left[   c^2 \hat{u}_i (ik_i \hat{\eta})^*  \right]\\
 	= &-\text{Re}\left[   c^2 \hat{u}_i^* ik_i \hat{\eta} \right]
\end{align}

%-------------------------------------------------%
\paragraph{Non-quadratic K.E. budget:}
\begin{align*}
  \pder{t}E_K^{NQ}(\mathbf{k},t) =& T_K^{NQ} + C_K^{NQ}
\end{align*}
Following equations \eqref{dtu0} and \eqref{dtM} for rate of change of velocity 
$\mathbf{u}$ and the displaced mass flux, $\mathbf M = \eta \mathbf{u}$ 
respectively,
 \begin{align*}
 \partial_t E_K^{NQ}(\mathbf{k},t)
 =& \frac{1}{2}\partial_t(\mathbf{\hat M}.\mathbf{\hat u^*})\\
 =& \frac{1}{2}\left[ \mathbf{\hat M} .\pder[\mathbf{\hat u^*}]{t} 
		      + \mathbf{\hat u^*}. \pder[\mathbf{\hat M}]{t}\right]\\
 =& \frac{1}{2}\left[-\hat{M}_i(\widehat{ u_j\partial_j u_i })^*
		      - c^2 \hat{M}_i (ik_i \hat{\eta})^*
		      - \hat{M}_i (\epsilon_{i3k} f \hat{u}_k)^* \right]\\
  &+\frac{1}{2}\left[ - \hat{u}_i^* \widehat{(u_j\partial_j M_i) }
		      - c^2\hat{u}_i^*ik_i\widehat{\eta\eta}/2
		      - \hat{u}_i^* (\epsilon_{i3k} f \hat{M}_k) 
		      - \hat{u_i}^* \widehat{(u_i \partial_j({u}_j+ 
M_j))}
			\right]
\end{align*}
Hence,
\begin{align}
 T_K^{NQ}= & -\frac{1}{2}\left[\hat{M}_i(\widehat{ u_j\partial_j u_i })^*
			      + \hat{u}_i^* \widehat{(u_j\partial_j M_i) }
			      + \hat{u_i}^* \widehat{(u_i \partial_j{M}_j)}
			      \right] \\
 T_{K,K}^{NQ,Q}= & -\frac{1}{2}\left[ \hat{u_i}^* \widehat{(u_i 
\partial_j{u}_j)}
			\right] \\
 T_{K,P}^{NQ,Q}= & -\frac{1}{2}\left[c^2 \hat{M}_i (ik_i \hat{\eta})^*
	      + c^2\hat{u}_i^*ik_i\widehat{\eta\eta} /2
	      \right]
\end{align}

\paragraphbf{Helmholtz decomposition of kinetic energy}

Furthermore, using Helmholtz decomposition to substitute for $\mathbf u$ in  
the expression 
for \emph{mean kinetic energy},

\begin{align}
\braket{ E_K} 
& = \frac{1}{2} \braket{ h\mathbf u.\mathbf u} \\
& = \frac{1}{2} \braket{h(\mathbf{u_r} + \mathbf{u_d} ).(\mathbf{u_r} + 
\mathbf{u_d} )}\\
& = \frac{1}{2}\braket{ h |\mathbf{u_r} | ^2} 
    -\braket{ \mathbf{u_r}.h\mathbf{u_d}}
    +  \frac{1}{2} \braket{ h | \mathbf{u_d}| ^2}
\end{align}
where, $h = 1 + \eta$ and, $\eta$ is the surface displacement of the shallow 
water layer. In the limit for small displacements i.e., $\eta \to 0$ or $ h \to 
1$, invoking orthogonality,

\begin{equation}
\lim_{h \to 1} \braket{E_K} = \braket{E_K^{Q}}
=  \frac{1}{2}\braket{|\mathbf{u_r}  | ^2 + | \mathbf{u_d} | ^2}
\end{equation}
The spectral equivalent of the above expression will be:
\begin{equation}
\braket{E_K^{Q} }
=  \frac{1}{2}\braket{ \mathbf{\hat{u}_r}.\mathbf{\hat{u}_r}^* + 
\mathbf{\hat{u}_d}.\mathbf{\hat{u}_d}^* }
\end{equation}

%---------------------------------------------------------%
\subsubsection{Available potential energy}
\paragraphbf{Expression for mean available potential energy}
\begin{align}
\braket{ E_P} 
& = \frac{c^2}{2} \braket{ \eta^2}
\end{align}
The spectral equivalent of which is:
\begin{align}
\braket{ E_P} 
& = \frac{c^2}{2} \braket{ \hat{\eta}\hat{\eta}^*}
\end{align}

\paragraphbf{A.P.E. budget}
Similar to equation \eqref{dtKE} we can form,
\begin{equation}\label{dtPE}
 \pder{t}E_P(\mathbf{k},t) = T_P+ C_P
\end{equation}
where, $T_P$ and $C_P$ represents the transfer and conversion spectral 
functions respectively.

Following \eqref{dth}, equation \eqref{dtPE} expands as:
\begin{align*}
 \pder{t}E_P(\mathbf{k},t)
 =& \frac{c^2}{2}\partial_t({\hat \eta}{\hat \eta^*})\\
 =& \frac{c^2}{2}\left[ \hat{\eta} .\pder[\hat{\eta}^*]{t}+ \hat 
\eta^*.\pder[\hat \eta]{t}\right]\\
 =& \frac{c^2}{2}\left[ - \hat{\eta} (ik_i\hat{ u_i } + ik_i\widehat{ 
      \eta u_i })^*  
	    - \hat{\eta}^* (ik_i\hat{ u_i } + ik_i\widehat{ 
      \eta u_i })  \right]\\
      =& -\text{Re}\left[c^2\hat{\eta} (ik_i\hat{ u_i } + ik_i\widehat{ 
      \eta u_i })^* \right]
\end{align*}
Thus,
\begin{align}
T_P= & -\text{Re}\left[c^2\hat{\eta} (ik_i\widehat{ 
      \eta u_i })^*  \right]\\
C_P=&  -\text{Re}\left[c^2\hat{\eta} (ik_i\hat{u_i })^*\right]\\
   =&  +\text{Re}\left[c^2\hat{\eta} ik_i\hat{u_i }^*\right]
\end{align}
thus, $C^Q_K = -C_P$ which, in turn, makes the assertion that  
equivalent conversion occurs between quadratic K.E. and A.P.E.
% 
% \begin{align*}
% 2\times T_K(\mathbf{k},t)
% & = \partial_t (\mathbf u.h\mathbf u)\\
% &= \partial_t \mathbf u .h\mathbf u  +|\mathbf u|^2 \partial_t h + h\mathbf u 
% .\partial_t \mathbf u \\
% & =   2 \left( -  \frac{\nabla|\mathbf u|^2}{2} - c^2 \nabla h - \zeta_a \times \mathbf u	\right).h\mathbf u - \left( |\mathbf u|^2 \nabla.(h \mathbf u )\right)\\ 
% & =   \left( -  h\frac{\nabla|\mathbf u|^3}{3} - c^2 \frac{\nabla h^2}{2} . \mathbf u \right)- \left( |\mathbf u|^2 \nabla.(h \mathbf u )\right) \\
% \end{align*}
% 
%

% \end{document}