### 2019-07-23 latex/glossary.tex
Move stuff

\makeglossaries

%% ------------
%% Definitions
%% ------------
\newglossaryentry{zeta}{name=$\zeta$, description={absolute vorticity}}
\newglossaryentry{kappa}{name=$\kappa$, description={magnitude of wavenumber}}
\newglossaryentry{sigma}{name=$\sigma$, description={frequency of IG waves, $\sqrt{f^2+(c\kappa)^2}$}}
\newglossaryentry{J}{name=$\mathbf{J}$, description={mass flux, $h\mathbf{u}$}}
\newglossaryentry{M}{name=$\mathbf{M}$, description={displaced mass flux, $\eta\mathbf{u}$}}
\newglossaryentry{W}{name=$\mathbf{W}$, description={derived variable vector}}
\newglossaryentry{U}{name=$\mathbf{U}$, description={primitive variable vector}}
\newglossaryentry{B}{name=$\mathbf{B}$, description={normal mode vector, dimension: $LT^{-1}$}}
\newglossaryentry{N}{name=$\mathbf{N}$, description={normal mode vector, dimension: $T^{-1}$}}
\newglossaryentry{Xn}{name=$X_n$, description={normalized eigenvector matrix}}
\newglossaryentry{P}{name=$P$, description={transformation matrix from $\mathbf{W}$ to $\mathbf{U}$}}
\newglossaryentry{Q}{name=$Q$, description={transformation matrix from $\mathbf{U}$ to $\mathbf{B}$}}

% \glossarystyle{mcolindex}
\glsaddall