### 2019-08-04 frontmatter.tex
Fix frontmatter

%-------------------------------------------------------------------------------
% Cover page
%-------------------------------------------------------------------------------
%
\makecoverpage

%-------------------------------------------------------------------------------
% Dedication (comment if none)
%-------------------------------------------------------------------------------
%
\dedication{
	% {\it ``You must unlearn what you have learnt.''}\\[20pt]
	% Master Yoda
	{\it ``Essentially, all models are wrong, but some are useful.''}\\[20pt]
	George E.~P.~Box
}

%-------------------------------------------------------------------------------
% Abstract (English)
%-------------------------------------------------------------------------------
%
% COMMENT: the abstract should be limited to 2000 characters and the keywords
% to 250 characters because otherwise they will not fit in the mailing sheet.
%
% TODO: abstract in english and swedish
\begin{abstract}
    \input{abstract}
	% Access points  must work. In fact, few cryptographers would disagree with
	% the emulation of robots, which embodies the technical principles of
	% complexity theory. Hyp, our new application for lossless information, is the
	% solution to all of these challenges.
	%         %
	%         \keywords{super-lasers, space-stations, The Force, cookies.}
	%         %
\end{abstract}

%-------------------------------------------------------------------------------
% Abstrakt (Swedish)
%-------------------------------------------------------------------------------
%
% COMMENT: the abstract should be limited to 2000 characters and the keywords
% to 250 characters because otherwise they will not fit in the mailing sheet.
%
\begin{abstrakt}
    \input{abstrakt}
	% \r{A}tkomstpunkter m\r{a}ste arbeta. I sj\"{a}lva verket skulle f\r{a}
	% kryptografer h\r{a}ller inte med emulering av robotar, som
	% f\"{o}rkroppsligar de tekniska principer av komplexitetsteori. Hyp, v\r{a}r
	% nya applikation f\"{o}r f\"{o}rlustfri information är l\"{o}sningen på alla
	% dessa utmaningar.
	%         %
	%         \nyckelord{super-lasrar, rymdstationer, Kraften, kakor.}
	%         %
\end{abstrakt}

%-------------------------------------------------------------------------------
% Preface page
%-------------------------------------------------------------------------------
%
\begin{preface}
    This thesis contains numerical, theoretical, experimental and
    technical studies aiming at advancing the understanding of stratified flows.
    Studies which motivated this thesis, along with theoretical
    details and research highlights of the present work are summarized in the
    first part. The second part contains five articles and one conference
    paper.
    The papers are adjusted to comply with the present thesis format for
    consistency, but their contents have not been altered as compared with their
    original counterparts.
\end{preface}

%-------------------------------------------------------------------------------
% Division of work between authors
%-------------------------------------------------------------------------------
%
\begin{divisionofwork}
	The main advisor for the thesis is Erik Lindborg (EL).
    Pierre Augier (PA) acts as co-advisor.

	\paperitem
        The code was developed by PA and Ashwin Vishnu
        Mohanan (AVM). The simulations and post-processing were performed
        by AVM. The theory was developed by PA and EL with input from AVM. A
        first version of the paper was written by PA. The final version was
        written by EL with input from AVM.
        %
	\paperitem
        The code development, the simulations and post-processing was
        handled by AVM. The paper was written by EL with input from
        AVM.
        %
    \paperitem
        The experiment was planned by the group from KTH together with the group from
        LEGI. The experimental set-up was put together by
        the group from LEGI. The experiment was carried out by the two groups
        together.
        The code for image-processing was developed by AVM, PA, Antoine
        Campagne (AC), Joel Sommeria (JS) and the code for processing
        probe data was developed by PA, RC and Diane Micard. The
        post-processing was performed by AVM, PA, AC, AS and JS.
        The paper was written by AC with input from PA, EL and AVM.
        %
    \paperitem
        The code was developed by PA, Cyrille Bonamy (CB) and AVM. The paper
        was authored by PA with input from AVM.
    \paperitem
        The code was developed by AVM, CB and PA.
        The code was benchmarked for performance by PA and AVM.
        The paper was written by AVM with inputs from PA and CB.
    \paperitem
        The code was developed by AVM, CB, Miguel Calpe Linares (MCL) and PA.
        The code was benchmarked for performance by PA and AVM.
        The paper was written by AVM with inputs from PA and CB.

\end{divisionofwork}

%TODO: fill in additional publications and conferences
%-------------------------------------------------------------------------------
% Additional publications (comment if none)
%-------------------------------------------------------------------------------
%
% \begin{otherpublications}
%         The following papers, although related, are not included in this thesis.
%
%   \paperitem%
%     {Anakin Skywalker \& Master Obi-Wan Kenobi}% Authors
%     {3639}% Year
%     {The light sabre: an elegant weapon for a more civilized age}% Title
%     {Jedi Journal of Weapons}% Journal
%     {33}% Volume
%     {2}% Number
%     {pp. 55--60}% Pages
%
% \end{otherpublications}

%-------------------------------------------------------------------------------
% Conferences (comment if none)
%-------------------------------------------------------------------------------
%
% \begin{conferences}
%         Part of the work in this thesis has been presented at the following
%         international conferences. The presenting author is underlined.
%
%   \conferenceitem%
%     {Anakin Skywalker, \underline{Darth Vader} \& Admiral Motti}% Authors
%     {3639}% Year
%     {Effects on breathing of underestimation of the power of The Force}% Title
%     {Intergalactic Conference on Weapons and Peace}% Conference
%     {Kamino}% Location
%
% \end{conferences}

%-------------------------------------------------------------------------------
% Table of contents
%-------------------------------------------------------------------------------
%
\tableofcontents