### 2019-08-05 templates/paper.tex

% Define title, author(s), affiliation and publishing status
%
\papertitle[Title] % Short title used in headlines (optional)
{%
  Long title% THE COMMENT SYMBOL AT THE END OF THIS LINE IS NEEDED
}%
%
\papertoctitle{Long title} % Title for toc
%
% Short authors used in headlines and List Of Papers
\paperauthor[A. Beta, G. Delta \& E. Phi]
{%
  Alpha Beta$^1$, Gamma Delta$^2$ and Epsilon Phi$^2$ % Short authors used in headlines and List Of Papers
}%
%
% (optional) Short authors used in List Of Papers
% \listpaperauthor[A. Beta, G. Delta \& E. Phi]
%
\paperaffiliation
{%
      $^1$ Linn\'e FLOW Centre, KTH Mechanics, S-100 44 Stockholm, Sweden \\
      $^2$ Ancient Rome University
}%
%
\paperjournal[Gal. Empire Pub.] % Short publish info used in List Of Papers
{%
	Galactic Empire Publications%
}%
%
\papervolume{42}%
%
\papernumber{2}%
%
\paperpages{1--10}%
%
\paperyear{3639}%
%
\papersummary%
{% Insert summary of the paper here (used in introduction)
    The implications of concurrent archetypes have been far-reaching and
pervasive. Given the current status of heterogeneous technology,
cyberinformaticians daringly desire the key unification of the Turing
machine and erasure coding. We explore new decentralized information,
which we call Tuna.

}%
%
\graphicspath{{paper1/}}%
%
%
%===============================================================================
%                            BEGIN PAPER
%===============================================================================
%
\begin{paper}

\makepapertitle

%------------------------------------------------------------------------------
% Abstract
%------------------------------------------------------------------------------
%
\begin{paperabstract}
    The implications of concurrent archetypes have been far-reaching and
pervasive. Given the current status of heterogeneous technology,
cyberinformaticians daringly desire the key unification of the Turing
machine and erasure coding. We explore new decentralized information,
which we call Tuna.

\end{paperabstract}

%------------------------------------------------------------------------------
% Article
%------------------------------------------------------------------------------
%
\input{paper1/article.tex}

%------------------------------------------------------------------------------
% Bibliography
%------------------------------------------------------------------------------
%
%\clearpage
%\bibliographystyle{jfm}
%\bibliography{thesis}
%\IfFileExists{paper1/paper.bbl}{\input{paper1/paper.bbl}}{}
\begin{refcontext}[sorting=nyt]
\printbibliography[heading=subbibliography]
\end{refcontext}
%===============================================================================
%                            END PAPER
%===============================================================================
\end{paper}